Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2023-06-16 analysis/paper_draft4.Rmd
moved Fig for 95%CI (ten strikes) to Supplement

Without historical strike data, there is a higher risk of underpowered and inflated estimates of attorney strike bias, because of the low number of strikes per trial: typically from 6 to  15 strikes per party, depending on the jurisdiction [@nscs-jury2018]. Yet, complete pooling of current and historical strike data may be affected by the degree of incompatibility between current and historical strike data. Such incompatibility may occur if missing historical strike is missing not at random due to incomplete or inaccessible jury-selection records [@Grosso2017; @Wright2018].

@@

We demonstrate this approach with simulations in which we vary the degree of incompatibility between current and historical trial data, and test how well we can "detect" attorney strike bias, i.e., whether the credible interval for the bias parameter excludes zero, given different values of the power prior's alpha parameter. Then, we present a software prototype that encodes the same approach with actual historical strike data from a convenience sample of criminal cases from one court. With that data, we use the prototype to illustrate how attorneys and others can use this approach to detect strike bias in real time during jury selection.

@@

```{r figbd9510, out.width = "95%", fig.cap = "Bias detection rates (proportion of simulations where 0 is excluded from the 95 percent credible interval of bias parameter) for simulations based on 10 strikes.", fig.align='center'}

load_path <- here::here("figures","pp10_95CI.png")
knitr::include_graphics(load_path)

@@

As we decrease the number of strikes each attorney has from fifteen to ten and six (Figures \@ref(fig:figbd9510) and \@ref(fig:figbd956) respectively), bias detection decreases. Bias detection is only high when $\alpha$ is high (close to 1); current trial bias is high; there is high compatibility between current and historical trials; and the number of historical trials is large.

>>>>

Without historical strike data, there is a higher risk of underpowered and inflated estimates of an attorney's strike bias, because of the low number of strikes per trial: typically from 6 to 15 strikes per party, depending on the jurisdiction [@nscs-jury2018]. Yet, complete pooling of current and historical strike data may be affected by the degree of incompatibility between current and historical strike data. Such incompatibility may occur if missing historical strike is missing not at random due to incomplete or inaccessible jury-selection records [@Grosso2017; @Wright2018].

@@

We demonstrate this approach with simulations in which we vary the degree of incompatibility between current and historical trial data, and test how well we can "detect" attorney strike bias, i.e., whether the credible interval for the bias parameter excludes zero, given different values of the power prior's alpha parameter. Then, we present a software prototype that encodes the same approach with actual historical strike data from a convenience sample of criminal cases from one court. With that data, we use the prototype to illustrate how attorneys and others can use this approach to detect strike bias during jury selection.

@@

@@

As we decrease the number of strikes each attorney has from fifteen to six (Figure \@ref(fig:figbd956)), bias detection decreases. (For the simulation results for ten strikes, see the Supplementary Material.) Bias detection is only high when $\alpha$ is high (close to 1); current trial bias is high; there is high compatibility between current and historical trials; and the number of historical trials is large.
